# Methodology

This is a study about the effects of vitamin C over the number of odontoblasts (cells responsible for tooth growth) of 60 guinea-pigs. If the number of odontoblasts grows, so does the teeth of the pigs.

Each pig received one of three different dosages of vitamin C: 0.5, 1 or 2 mg/day.

Each pig was treated by one of two different delivery methods: via orange juice or via ascorbic acid (VC)

We are going to test some hypothesis about the vitamin C influence over tooth growth.

There is not a control group.

# Summary of the data

Here is a brief summary of data:

```{r}
dados <- ToothGrowth
str(dados)
```


```{r}
summary(dados)
```

```{r fig.cap="Tooth Growth data", cache=TRUE}
plt <- ggplot(dados, aes(x=dose, y=len, group=supp, color=supp))
plt <- plt + geom_point() + 
        theme(plot.title = element_text(hjust = 0.5)) +
        labs(title="Tooth Growth", y="Growth", x="Dose")

plt
```


OJ means Orange Juice while VC means ascorbic acid.


# Hypothesis testing

## Vitamin C Dosis

Since the pigs were treated with different doses of Vitamin C, the first question we need to address is whether the amount of vitamin C administered to each pig does any difference in their tooth growth.

The different doses were:

```{r}
unique(dados$dose)
```

So, since there is not a control group, lets test the hypothesis that the doses matter: is the growth associated with the 2.0 mg doses greater than the growth associated with 0.5 doses?

$$H_0: \mu_{0.5 mg} = \mu_{2.0 mg} \leftrightarrow H_\alpha: \mu_{2.0 mg} > \mu_{0.5 mg} $$

Lets split the 2 groups and test the hypothesis through a one-sided T-test 

```{r echo=TRUE}
g05 <- subset(dados$len, dados$dose == 0.5)
g2 <- subset(dados$len, dados$dose == 2)

t.test(g2, g05, alternative = "greater", paired = FALSE, var.equal = FALSE)
```

Notice the very small p-value. 

Just for comparison, lets do this test *by hand* through its confidence interval:

```{r echo=TRUE}
n2 <- length(g2)
n05 <- length(g05)
sd2 <- sd(g2)
sd05 <- sd(g05)
se1 <- sqrt(((n2-1)*sd2^2 + (n05-1)*sd05^2) / (n2 + n05 -2))
se2 <- sqrt(1/n2 + 1/n05)
se <- se1*se2
ci <- mean(g2) - mean(g05) + c(-1,1)*qt(.95,df = (n2+n05 -2))*se
ci <- round(ci,2)
ci
```
Notice that 0 isn't in the confidence interval, so $\mu_{0.5 mg} - \mu_{2.0 mg} \neq 0$, ever!
The true difference of means will always be between `r ci` mm.


We can infer by this 95% test that $H_0$ must be rejected and therefore, **a larger dosis results in larger tooth growth!**

## Vitamin C Delivery Method

What is the best delivery method? 

The delivery methods are:

```{r}
unique(dados$supp)
```

OJ means Orange Juice while VC means ascorbic acid.

```{r fig.cap="Tooth Growth by delivery method", cache=TRUE}
plt <- ggplot(dados, aes(x=supp, y=len, group=factor(dose), color=factor(dose)))
plt <- plt + geom_point() + 
        theme(plot.title = element_text(hjust = 0.5)) +
        labs(title="Tooth Growth", y="Growth", x="Delivery Method")

plt
```


Lets test the null hypothesis that the deliver methods yields the same results:

$$H_0: \mu_{juice} = \mu_{acid} \leftrightarrow H_\alpha: \mu_{juice} \neq \mu_{acid}$$ 

The first thing to do is to separate the data between the two delivery methods and then run a two-sided T-test on the groups.


```{r echo=TRUE}

juicers <- subset(dados$len, dados$supp=="OJ")
trippin <- subset(dados$len, dados$supp=="VC")

t.test(juicers, trippin, paired = FALSE, var.equal = FALSE, alternative = "two.sided")
```

Interesting enough, the T-test yields a confidence interval and a p-value that led us to **fail to reject $H_0$.**

Therefore, **the two delivery methods produces the same results**


# Results

As we can see through the T-tests performed, we can infer that:

1. a larger doses results in larger tooth growth
1. the two delivery methods produces the same results

